Active Learning for SAT-Solver Benchmarking

### 2022-10-04 paper/paper.tex
Add more details to some comments in paper
Also, make a few suggestions for further related work.
Add some related work
... in all four subsections of that section. Not many relevant works found.

We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
%JB: is there any further behavior we can deduce about a solver scoring than a performance ranking? similar behavior could be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.
%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
%JB: we can probably cut this part a bit, as we discuss weakness of individual approaches later anyway
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
%JB: as commented above, we should either remove this criterion or show it in the table

\subsubsection{Traditional benchmark-set-selection.}
%JB: I think we should do some further research on static benchmark-set selection approaches (maybe in SMT, CP, etc.)
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.

@@

\begin{table}[tbp]
  \centering
  \caption{
  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
  }
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

@@

While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
%JB: again, I think there are too many details for describing this single approach
Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

@@

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment
Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
%JB: this general motivation should go there where linearity is introduced as criterion, not here
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

@@

Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
%JB: we should conduct some further research if there are other approaches using AL for SAT (or SMT, CP, etc.) benchmarking
This goal is, however, orthogonal to ours.
While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark set.

@@

@@

\section{Active Learning for SAT-Solver Benchmarking}
\label{sec:main}
Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
We \emph{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
Refer to the next sections for more details.

@@

  $\operatorname{score} \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
  %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"? or return a tuple with all estimated scores (one per solver)?; also, one could expect to get a benchmark set (see title of algorithm), but as instance selection is solver-specific, this is not case

  \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
  %JB: we should formally specify the tuple rather than only describing it
\end{algorithm}

@@

Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes does not work well in practice.
%JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Sec.~\ref{sec:eval}).
The following paragraphs motivate why it makes sense to discretize runtimes.
%JB: we probably should cut this a bit

We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
In our experiments, this function is a clustering algorithm.
Note that this transformation is done per instance to adapt to different instance difficulties. 
Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
  %JB: does the "-1" provide any benefits, or does it only shift scores?
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
%JB: as "$\operatorname{score}$" is not visible, we should use some other way of highlighting

@@

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
This means that we do not lose much discriminatory power using discrete runtime labels.
The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
%JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
%JB: at the same time with what? with the other benefits?

\subsection{Instance Sampling}
\label{sec:main2}
Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.

@@

    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)

>>>>

Note that we know the runtimes of all solvers, except the new one, on all instances.
There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
%JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking? "similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
%JB: we should rename "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
%JB: for what do we measure cost here?
% (1) solving new-solver-problem once? -> apart from more runtimes to be acquired beforehand, is running our framework affected by increasing the set of existing solvers, and how does this depend on the configuration of our framework, i.e., sampling/selection/stopping? (e.g., Wilcoxon stopping criterion iterates over set of existing solvers, showing linear complexity, while fixed-subset stopping is independent from number of existing solvers)
% (2) solving new-solver-problem for each solver out of a set of solvers? -> is a rather theoretical scenario, because our algorithm itself assumes we know runtimes of all but one solver, so in practice we cannot throw in a bunch of new solvers and get a ranking over all of them
We avoid comparing pairs of solvers, which would scale quadratically.
%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if binary-search-/recursive-sorting-like procedure used
Since a new solver idea is rarely best at first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

@@

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
%JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
%JB: as commented above, we should either remove this criterion or show it in the table

\subsubsection{Traditional benchmark-set-selection.}
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

Benchmark-set selection is an important issue for solver competitions, e.g., the SAT Competition.
In such competitions, the organizers define the rules for composing the corresponding benchmarks.
Balint et al.~\cite{balint2015overview} provide an overview of benchmark-selection criteria in different solver competitions.
%JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al. (2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
M{\i}s{\i}r~\cite{misir2021benchmark} presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances.
The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed criteria
Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.

@@

\begin{table}[tbp]
  \centering
  \caption{
  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
  }
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

@@

While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
If the new solver does not belong to this category (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
%JB: "this category" = category of most promising solvers? not 100% clear
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
The main difference is that Matricon et. al. compare a solver to another one, while we compare to a set of existing solvers.
To the best of our knowledge, we are the first to tackle the latter problem.
%JB: again, I think there are too many details for describing this single approach

Given a pair of SAT solvers, Matricon et. al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

@@

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
%JB: maybe rather log-linear than quadratic, see a prior comment in intro

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure if this dilemma arises here as well; dilemma is typical for searching one optimal solution in continuous space, where there is an incentive to exploit neighborhood in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well

@@

Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
This goal is, however, orthogonal to ours.
While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark set.

@@

In the field of SAT-solving, Volpato and Guangyan~\cite{volpato2019active} use uncertainty-sampling-based AL to learn an instance-specific algorithm selector.
Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.

@@

\section{Active Learning for SAT-Solver Benchmarking}
\label{sec:main}
Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
%JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
We \emph{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
Refer to the next sections for more details.

@@

  $\operatorname{score} \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
  %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"? or return a tuple with all estimated scores (one per solver?); also, one could expect to get a benchmark set as result (see title of algorithm), but since instance selection is solver-specific, this is not case (it's rather a benchmarking schema)

  \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
  %JB: we should formally specify the tuple rather than only describing it in words
\end{algorithm}

@@

Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes does not work well in practice.
%JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Sec.~\ref{sec:eval}).
The following paragraphs motivate why it makes sense to discretize runtimes.
%JB: we should probably cut this a bit

We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
In our experiments, this function is a clustering algorithm.
Note that this transformation is done per instance to adapt to different instance difficulties. 
Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
  %JB: does the "-1" provide any benefits, or does it only shift scores?
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
%JB: as "$\operatorname{score}$" is not visibly different to rest of text, we should use some other way of highlighting

@@

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
This means that we do not lose much discriminatory power using discrete runtime labels.
The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
%JB: thus, we should reorder this section to mention argument "significant differenc"e (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
%JB: at the same time with what? with the other benefits?

\subsection{Instance Sampling}
\label{sec:main2}
Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, so distinction between the two categories of approaches not entirely clear
Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.

@@

    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)?
